COVID-19 Closures Negatively Impact Restaurant Revenue and Employment

### 2021-02-23 outputs/paper/paper.Rmd

```{r treatmentcontrolgroups, echo = FALSE, message = FALSE, warning = FALSE}
# format table to show LHAs that were selected for treatment and control groups
treatment_control_groups %>%
  knitr::kable(caption = "Primary sample of Local Health Authorities from strata based on population size",
               col.names = c("Group", 
                             "Large Population LHAs", 
                             "Medium Population LHAs", 
                             "Small Population LHAs"),
               booktabs = TRUE) %>%
  kableExtra::kable_styling() %>%
  collapse_rows() %>%
  column_spec(1, color = "#972D15")
```

@@

```{r sampleallunitsdata, echo = FALSE, message = FALSE, warning = FALSE}
# format table to show a few rows of the complete list of restaurants in treatment and control groups
all_units_data %>%
  slice(101,1100,2101,3101,4100,5100,6100,7100,8100,9101) %>%
  knitr::kable(caption = "Example segment of the secondary sampling frame",
               booktabs = TRUE) %>%
  column_spec(1, width = "14em") %>%
  column_spec(2, width = "14em") %>%
  column_spec(3, width = "6em") %>%
  column_spec(4, width = "6em") %>%
  kableExtra::kable_styling(latex_options = c("scale_down")) %>%
  row_spec(0, color = "#972D15")
```

@@

# set the colors for the milestone events
status_levels <- c("Survey 1", "Treatment", "Survey 2")
status_colors <- c("#D8B70A", "#972D15", "#5BBCD6")

@@

# plot the data
ggplot(df,aes(x = date, y = 0, col = status, label = milestone)) +
  labs(col = "") +
  scale_color_manual(values = status_colors, labels = status_levels, drop = FALSE) +
  theme_classic() +
  # Plot horizontal black line for timeline
  geom_hline(yintercept = 0, color = "black", size = 0.3) +
  # Plot vertical segment lines for milestones
  geom_segment(data = df[df$month_count == 1,], aes(y = position, yend = 0, xend = date), color = 'black', size = 0.2) +
  # plot colored lines for ongoing events
  geom_segment(aes(x = as_date("2021-05-01"), xend = as_date("2021-05-31"), y = 0, yend = 0), color = "#D8B70A", size = 1) +
  geom_segment(aes(x = as_date("2021-07-01"), xend = as_date("2021-07-31"), y = 0, yend = 0, ), color = "#5BBCD6", size = 1) +
  # Plot scatter points at zero and date
  geom_point(aes(y = 0), size = 3) +
  # Don't show axes, appropriately position legend
  theme(axis.line.y=element_blank(),
                 axis.text.y = element_blank(),
                 axis.title.x = element_blank(),
                 axis.title.y = element_blank(),
                 axis.ticks.y = element_blank(),
                 axis.text.x = element_blank(),
                 axis.ticks.x = element_blank(),
                 axis.line.x = element_blank(),
                 legend.position = "bottom"
                ) +
  # Show text for each month
  geom_text(data = month_df, aes(x = month_date_range, y = -0.05, label = month_format), size = 3, vjust = 0.5, color = 'black') +
  # Show text for each milestone
  geom_text(aes(x = c(as_date("2021-06-03"), as_date("2021-06-26"), as_date("2021-06-25"), as_date("2021-07-01"), as_date("2021-07-15"), as_date("2021-08-03"), as_date("2021-08-27")),
                y = text_position, label = milestone),size=3) +
  # add labels for the ongoing events
  geom_text(aes(x = as_date("2021-05-16"), y = 0.1, label = "Study period of Survey 1"), size = 3, color = "#D8B70A") +
  geom_text(aes(x = as_date("2021-07-16"), y = 0.1, label = "Study period of Survey 2"), size = 3, color = "#5BBCD6")
```

@@

# put the results in a table
  knitr::kable(caption = "Summary statistics for treatment and control baselines",
               booktabs = TRUE, linesep = "",
               col.names = c("", "", "Treatment", "Control")) %>%
  kableExtra::kable_styling(latex_options = "striped", stripe_index =c(1,5:7,11:13,17:19), stripe_color = "#fbf7e6") %>%
  column_spec(1, color = "#972D15")
```

```{r baselinerevenue, fig.cap="Revenue distribution for treatment and control baselines, from Survey 1", fig.width=10, fig.height=4, echo = FALSE, message = FALSE, warning = FALSE}
# plot survey 1 revenue distribution, separating treatment and control
ggplot(survey_1, aes(x = revenue / 1000, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 5) +
  theme_minimal() +
  scale_x_continuous(limits = c(0, 250), breaks = c(0, 50, 100, 150, 200, 250, 300)) +
  labs(fill = NULL,
       x = "Revenue in May 2021 (in thousands of Canadian dollars)",
       y = "Number of restaurants",
       title = "Revenue distributions were similar in the treatment and control samples") +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))
```

@@

# make the full time employment count plot
p1 <- ggplot(survey_1, aes(x = pt, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 1, show.legend = FALSE) +
  coord_flip() +
  xlim(0,35) +
  scale_y_continuous(limits = c(0,135), breaks = c(0,25,50,75,100,125,150)) +
  theme_minimal() +
  labs(x = "Number of employees in a restaurant",
       y = "Number of restaurants",
       title = "Part-time employees") +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))

# make the part time employment count plot
p2 <- ggplot(survey_1, aes(x = ft, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 1) +
  coord_flip() +
  xlim(0,35) +
  scale_y_continuous(limits = c(0,75), breaks = c(0,25,50,75)) +
  theme_minimal() +
  labs(fill = NULL,
       x = NULL,
       y = "Number of restaurants",
       title = "Full-time employees")  +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))

@@

# put the results in a table
  knitr::kable(caption = "Summary statistics for treatment and control groups post-treatment",
               booktabs = TRUE, linesep = "",
               col.names = c("", "", "Treatment", "Control")) %>%
   kableExtra::kable_styling(latex_options = "striped", stripe_index =c(1,5:7,11:13,17:19), stripe_color = "#fbf7e6") %>%
  column_spec(1, color = "#972D15")
```

@@

```{r postrevenue, fig.cap="Revenue distribution for treatment and control groups, from Survey 2", fig.width=10, fig.height=4, echo = FALSE, message = FALSE, warning = FALSE}
# plot survey 2 revenue distribution, separating treatment and control
ggplot(survey_2, aes(x = revenue / 1000, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 5) +
  theme_minimal() +
  scale_x_continuous(breaks = c(0, 50, 100, 150, 200, 250, 300)) +
  labs(fill = NULL,
       x = "Revenue in July 2021 (in thousands of Canadian dollars)",
       y = "Number of restaurants",
       title = "The revenue distribution was lower for the treatment sample") +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))
```

@@

# make the full time employment count plot
p1 <- ggplot(survey_2, aes(x = pt, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 1, show.legend = FALSE) +
  coord_flip() +
  xlim(0,35) +
  scale_y_continuous(limits = c(0,135), breaks = c(0,25,50,75,100,125)) +
  theme_minimal() +
  labs(x = "Number of employees in a restaurant",
       y = "Number of restaurants",
       title = "Part-time employees") +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))

# make the part time employment count plot
p2 <- ggplot(survey_2, aes(x = ft, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 1) +
  coord_flip() +
  xlim(0,35) +
  scale_y_continuous(limits = c(0,75), breaks = c(0,25,50,75)) +
  theme_minimal() +
  labs(fill = NULL,
       x = NULL,
       y = "Number of restaurants",
       title = "Full-time employees") +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))

@@

# create plot using summary data of means, original dataframe, and the split violin function above:
ggplot() +
  geom_split_violin(data = (survey_2 %>% filter(service_type != "NA")), aes(x = service_type, y = revenue / 1000, fill = group), color = NA) +
  # add lines to indicate means of the distributions
  geom_errorbar(data = means, 
                aes(x = service_type, y = mean, ymax = mean, ymin = mean, fill = group), 
                position = "dodge",
                color = "#972D15") +
  labs(fill = NULL,
       x = "Service type provided",
       title = "Revenues were more heavily impacted for dine-in only establishments",
       y = "Revenue (in thousands of Canadian dollars)") +
  scale_x_discrete(labels = c( "Dine-in only", "Both dine-in and take-out", "Take-out only")) +
  theme_minimal() +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))
```

@@

# create plot using summary data of means, original dataframe, and the split violin function above:  
ggplot() +
  geom_split_violin(data = data, aes(x = minority, y = revenue / 1000, fill = group), color = NA) +
  # add lines to indicate means of the distributions
  geom_errorbar(data = means,
                aes(x = minority, y = mean, ymax = mean, ymin = mean, fill = group),
                position = "dodge",
                color = "#972D15") +
  labs(fill = NULL,
       x = "Racial and ethnic minorities",
       title = "Revenues were more heavily impacted for minority-owned businesses",
       y = "Revenue (in thousands of Canadian dollars)") +
  theme_minimal() +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))
```

@@

# make tibble containing above numbers
tibble(Group = c("Treatment", "Control", "Treatment", "Control", "Treatment", "Control", "Treatment", "Control", "Treatment", "Control", "Treatment", "Control"), 
        Response = c("Complete", "Complete", "Partial", "Partial", "No response", "No response", "Complete", "Complete", "Partial", "Partial", "No response", "No response"), 
        Survey = c("Survey 1", "Survey 1","Survey 1","Survey 1","Survey 1","Survey 1", "Survey 2", "Survey 2", "Survey 2", "Survey 2", "Survey 2", "Survey 2"), 
        Count = c(num_treat_complete_survey1, num_control_complete_survey1, num_treat_partial_survey1, num_control_partial_survey1, num_treat_nonre_survey1, num_control_nonresurvey1,
          num_treat_complete_survey2, num_control_complete_survey2, num_treat_partial_survey2, num_control_partial_survey2, num_treat_nonre_survey2, num_control_nonresurvey2)) %>%
  # plot the data
  ggplot(aes(x = Response, y = Count, fill = Group)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "More than half of surveyed restaurants did not respond") +
  facet_wrap(~Survey) +
  theme_minimal() +
  scale_fill_manual(values = c("#5BBCD6", "#D8B70A"))
```

@@

# display survey budget in a table
survey_budget %>% 
  kable(digits = 2,
        caption = "2021 Ontario Restaurant Survey Budget") %>% 
  kable_styling() %>% 
  row_spec(0, bold = T, color = "white", background = "#972D15") %>% 
  row_spec(4, bold = T) %>%
  kableExtra::kable_styling(latex_options = c("hold_position"))
```

>>>>

```{r treatmentcontrolgroups, echo = FALSE, message = FALSE, warning = FALSE}
# format table to show LHAs that were selected for treatment and control groups
treatment_control_groups %>%
  knitr::kable(caption = "Primary sample of Local Health Authorities from strata based on population size",
               col.names = c("Group", 
                             "Large Population LHAs", 
                             "Medium Population LHAs", 
                             "Small Population LHAs"),
               booktabs = TRUE) %>%
  kableExtra::kable_styling() %>%
  collapse_rows() %>%
  column_spec(1, color = "#cd6423")
```

@@

```{r sampleallunitsdata, echo = FALSE, message = FALSE, warning = FALSE}
# format table to show a few rows of the complete list of restaurants in treatment and control groups
all_units_data %>%
  slice(101,1100,2101,3101,4100,5100,6100,7100,8100,9101) %>%
  knitr::kable(caption = "Example segment of the secondary sampling frame",
               booktabs = TRUE) %>%
  column_spec(1, width = "14em") %>%
  column_spec(2, width = "14em") %>%
  column_spec(3, width = "6em") %>%
  column_spec(4, width = "6em") %>%
  kableExtra::kable_styling(latex_options = c("scale_down")) %>%
  row_spec(0, color = "#cd6423")
```

@@

# set the colors for the milestone events
status_levels <- c("Survey 1", "Treatment", "Survey 2")
status_colors <- c("#D8B70A", "#cd6423", "#84b7d0")

@@

# plot the data
ggplot(df,aes(x = date, y = 0, col = status, label = milestone)) +
  labs(col = "") +
  scale_color_manual(values = status_colors, labels = status_levels, drop = FALSE) +
  theme_classic() +
  # Plot horizontal black line for timeline
  geom_hline(yintercept = 0, color = "black", size = 0.3) +
  # Plot vertical segment lines for milestones
  geom_segment(data = df[df$month_count == 1,], aes(y = position, yend = 0, xend = date), color = 'black', size = 0.2) +
  # plot colored lines for ongoing events
  geom_segment(aes(x = as_date("2021-05-01"), xend = as_date("2021-05-31"), y = 0, yend = 0), color = "#D8B70A", size = 1) +
  geom_segment(aes(x = as_date("2021-07-01"), xend = as_date("2021-07-31"), y = 0, yend = 0, ), color = "#84b7d0", size = 1) +
  # Plot scatter points at zero and date
  geom_point(aes(y = 0), size = 3) +
  # Don't show axes, appropriately position legend
  theme(axis.line.y=element_blank(),
                 axis.text.y = element_blank(),
                 axis.title.x = element_blank(),
                 axis.title.y = element_blank(),
                 axis.ticks.y = element_blank(),
                 axis.text.x = element_blank(),
                 axis.ticks.x = element_blank(),
                 axis.line.x = element_blank(),
                 legend.position = "bottom"
                ) +
  # Show text for each month
  geom_text(data = month_df, aes(x = month_date_range, y = -0.05, label = month_format), size = 3, vjust = 0.5, color = 'black') +
  # Show text for each milestone
  geom_text(aes(x = c(as_date("2021-06-03"), as_date("2021-06-26"), as_date("2021-06-25"), as_date("2021-07-01"), as_date("2021-07-15"), as_date("2021-08-03"), as_date("2021-08-27")),
                y = text_position, label = milestone),size=3) +
  # add labels for the ongoing events
  geom_text(aes(x = as_date("2021-05-16"), y = 0.1, label = "Study period of Survey 1"), size = 3, color = "#D8B70A") +
  geom_text(aes(x = as_date("2021-07-16"), y = 0.1, label = "Study period of Survey 2"), size = 3, color = "#84b7d0")
```

@@

# put the results in a table
  knitr::kable(caption = "Summary statistics for treatment and control baselines",
               booktabs = TRUE, linesep = "",
               col.names = c("", "", "Treatment", "Control")) %>%
  kableExtra::kable_styling(latex_options = "striped", stripe_index =c(1,5:7,11:13,17:19), stripe_color = "#f9f9f9") %>%
  column_spec(1, color = "#cd6423")
```

```{r baselinerevenue, fig.cap="Revenue distribution for treatment and control baselines, from Survey 1", fig.width=10, fig.height=4, echo = FALSE, message = FALSE, warning = FALSE}
# plot survey 1 revenue distribution, separating treatment and control
ggplot(survey_1, aes(x = revenue / 1000, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 5) +
  theme_minimal() +
  scale_x_continuous(limits = c(0, 250), breaks = c(0, 50, 100, 150, 200, 250, 300)) +
  labs(fill = NULL,
       x = "Revenue in May 2021 (in thousands of Canadian dollars)",
       y = "Number of restaurants",
       title = "Revenue distributions were similar in the treatment and control samples") +
  scale_fill_manual(values = c("#84b7d0", "#D8B70A"))
```

@@

# make the full time employment count plot
p1 <- ggplot(survey_1, aes(x = pt, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 1, show.legend = FALSE) +
  coord_flip() +
  xlim(0,35) +
  scale_y_continuous(limits = c(0,135), breaks = c(0,25,50,75,100,125,150)) +
  theme_minimal() +
  labs(x = "Number of employees in a restaurant",
       y = "Number of restaurants",
       title = "Part-time employees") +
  scale_fill_manual(values = c("#84b7d0", "#D8B70A"))

# make the part time employment count plot
p2 <- ggplot(survey_1, aes(x = ft, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 1) +
  coord_flip() +
  xlim(0,35) +
  scale_y_continuous(limits = c(0,75), breaks = c(0,25,50,75)) +
  theme_minimal() +
  labs(fill = NULL,
       x = NULL,
       y = "Number of restaurants",
       title = "Full-time employees")  +
  scale_fill_manual(values = c("#84b7d0", "#D8B70A"))

@@

# put the results in a table
  knitr::kable(caption = "Summary statistics for treatment and control groups post-treatment",
               booktabs = TRUE, linesep = "",
               col.names = c("", "", "Treatment", "Control")) %>%
   kableExtra::kable_styling(latex_options = "striped", stripe_index =c(1,5:7,11:13,17:19), stripe_color = "#f9f9f9") %>%
  column_spec(1, color = "#cd6423")
```

@@

```{r postrevenue, fig.cap="Revenue distribution for treatment and control groups, from Survey 2", fig.width=10, fig.height=4, echo = FALSE, message = FALSE, warning = FALSE}
# plot survey 2 revenue distribution, separating treatment and control
ggplot(survey_2, aes(x = revenue / 1000, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 5) +
  theme_minimal() +
  scale_x_continuous(breaks = c(0, 50, 100, 150, 200, 250, 300)) +
  labs(fill = NULL,
       x = "Revenue in July 2021 (in thousands of Canadian dollars)",
       y = "Number of restaurants",
       title = "The revenue distribution was lower for the treatment sample") +
  scale_fill_manual(values = c("#84b7d0", "#D8B70A"))
```

@@

# make the full time employment count plot
p1 <- ggplot(survey_2, aes(x = pt, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 1, show.legend = FALSE) +
  coord_flip() +
  xlim(0,35) +
  scale_y_continuous(limits = c(0,135), breaks = c(0,25,50,75,100,125)) +
  theme_minimal() +
  labs(x = "Number of employees in a restaurant",
       y = "Number of restaurants",
       title = "Part-time employees") +
  scale_fill_manual(values = c("#84b7d0", "#D8B70A"))

# make the part time employment count plot
p2 <- ggplot(survey_2, aes(x = ft, fill = group)) +                       
  geom_histogram(position = "dodge", binwidth = 1) +
  coord_flip() +
  xlim(0,35) +
  scale_y_continuous(limits = c(0,75), breaks = c(0,25,50,75)) +
  theme_minimal() +
  labs(fill = NULL,
       x = NULL,
       y = "Number of restaurants",
       title = "Full-time employees") +
  scale_fill_manual(values = c("#84b7d0", "#D8B70A"))

@@

# create plot using summary data of means, original dataframe, and the split violin function above:
ggplot() +
  geom_split_violin(data = (survey_2 %>% filter(service_type != "NA")), aes(x = service_type, y = revenue / 1000, fill = group), color = NA) +
  # add lines to indicate means of the distributions
  geom_errorbar(data = means, 
                aes(x = service_type, y = mean, ymax = mean, ymin = mean, fill = group), 
                position = "dodge",
                color = "#cd6423") +
  labs(fill = NULL,
       x = "Service type provided",
       title = "Revenues were more heavily impacted for dine-in only establishments",
       y = "Revenue (in thousands of Canadian dollars)") +
  scale_x_discrete(labels = c( "Dine-in only", "Both dine-in and take-out", "Take-out only")) +
  theme_minimal() +
  scale_fill_manual(values = c("#84b7d0", "#D8B70A"))
```

@@

# create plot using summary data of means, original dataframe, and the split violin function above:  
ggplot() +
  geom_split_violin(data = data, aes(x = minority, y = revenue / 1000, fill = group), color = NA) +
  # add lines to indicate means of the distributions
  geom_errorbar(data = means,
                aes(x = minority, y = mean, ymax = mean, ymin = mean, fill = group),
                position = "dodge",
                color = "#cd6423") +
  labs(fill = NULL,
       x = "Racial and ethnic minorities",
       title = "Revenues were more heavily impacted for minority-owned businesses",
       y = "Revenue (in thousands of Canadian dollars)") +
  theme_minimal() +
  scale_fill_manual(values = c("#84b7d0", "#D8B70A"))
```

@@

# make tibble containing above numbers
tibble(Group = c("Treatment", "Control", "Treatment", "Control", "Treatment", "Control", "Treatment", "Control", "Treatment", "Control", "Treatment", "Control"), 
        Response = c("Complete", "Complete", "Partial", "Partial", "No response", "No response", "Complete", "Complete", "Partial", "Partial", "No response", "No response"), 
        Survey = c("Survey 1", "Survey 1","Survey 1","Survey 1","Survey 1","Survey 1", "Survey 2", "Survey 2", "Survey 2", "Survey 2", "Survey 2", "Survey 2"), 
        Count = c(num_treat_complete_survey1, num_control_complete_survey1, num_treat_partial_survey1, num_control_partial_survey1, num_treat_nonre_survey1, num_control_nonresurvey1,
          num_treat_complete_survey2, num_control_complete_survey2, num_treat_partial_survey2, num_control_partial_survey2, num_treat_nonre_survey2, num_control_nonresurvey2)) %>%
  # plot the data
  ggplot(aes(x = Response, y = Count, fill = Group)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "More than half of surveyed restaurants did not respond") +
  facet_wrap(~Survey) +
  theme_minimal() +
  scale_fill_manual(values = c("#84b7d0", "#D8B70A"))
```

@@

# display survey budget in a table
survey_budget %>% 
  kable(digits = 2,
        caption = "2021 Ontario Restaurant Survey Budget") %>% 
  kable_styling() %>% 
  row_spec(0, bold = T, color = "white", background = "#cd6423") %>% 
  row_spec(4, bold = T) %>%
  kableExtra::kable_styling(latex_options = c("hold_position"))
```
